---
output:
  md_document:
    variant: markdown_github
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
stopifnot(require(knitr))
options(width = 90)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "png",
  dpi = 150,
  fig.asp = 0.8,
  fig.width = 5,
  out.width = "60%",
  fig.align = "center"
)
library(thurstonianIRT)
set.seed(1234)
```

# thurstonianIRT

[![Build Status](https://travis-ci.org/paul-buerkner/thurstonianIRT.svg?branch=master)](https://travis-ci.org/paul-buerkner/thurstonianIRT)
[![CRAN Version](http://www.r-pkg.org/badges/version/thurstonianIRT)](https://cran.r-project.org/package=thurstonianIRT)

## Overview 

The **thurstonianIRT** package allows to fit various models from Item Response
Theory (IRT) for forced-choice questionnaires, most notably the Thurstonian IRT
model originally proposed by (Brown & Maydeu-Olivares, 2011). IRT in general
comes with several advantages over classical test theory, for instance, the
ability to model varying item difficulties as well as item factor loadings on
the participants' traits they are supposed to measure. Moreover, if multiple
traits are modeled at the same time, their correlation can be incorporated into
an IRT model to improve the overall estimation accuracy. The key characteristic
of forced-choice questionnaires is that participants cannot endorse all items at
the same time and instead have to make a comparative judgment between two or
more items. Such a format comes with the hope of providing more valid inference
in situation where participants have motivation to not answer honestly (e.g., in
personnel selection), but instead respond in a way that appears favorable in the
given situation. Whether forced-choice questionnaires and the corresponding IRT
models live up to this hope remains a topic of debate (e.g., see Bürkner,
Schulte, & Holling, 2019) but it is in any case necessary to provide software
for fitting these statistical models both for practical and research purposes.

In the original formulation, the Thurstonian IRT model works on dichotomous
pairwise comparisons and models the probability of endorsing one versus the
other item. This probability depends on parameters related to the items under
comparison as well as on parameters related to the participants' latent traits
which are assumed to be measured by the items. For more details see Brown and
Maydeu-Olivares (2011), Brown and Maydeu-Olivares (2012), and Bürkner et al.
(2019).

## How to use thurstonianIRT

```{r}
library(thurstonianIRT)
```

As a simple example consider a data set of 4 blocks each
containing 3 items (i.e., triplets) answered by 200 participants.

```{r}
data("triplets")
head(triplets)
```

In the data set, a 1 indicates that the first item has been selected over the
second item while a 0 indicates that the second items has been selected
over the first item. In order to fit a Thurstonian IRT model on this data,
we have to tell **thurstonianIRT** about the block structure of the items,
the traits on which the items load, and the sign of these loadings, that is,
whether items have been inverted. For the present data, we specify
this as follows:

```{r}
blocks <-
  set_block(c("i1", "i2", "i3"), traits = c("t1", "t2", "t3"),
            signs = c(1, 1, 1)) +
  set_block(c("i4", "i5", "i6"), traits = c("t1", "t2", "t3"),
            signs = c(-1, 1, 1)) +
  set_block(c("i7", "i8", "i9"), traits = c("t1", "t2", "t3"),
            signs = c(1, 1, -1)) +
  set_block(c("i10", "i11", "i12"), traits = c("t1", "t2", "t3"),
            signs = c(1, -1, 1))
```

Next, we transform the data into a format that **thurstonianIRT** understands.

```{r}
triplets_long <- make_TIRT_data(
  data = triplets, blocks = blocks, direction = "larger",
  format = "pairwise", family = "bernoulli", range = c(0, 1)
)
head(triplets_long)
```

Finally, we can fit the model using several model fitting engines. Currently
supported are [Stan](https://mc-stan.org/), [lavaan](http://lavaan.ugent.be/),
and [Mplus](http://www.statmodel.com/). Here, we choose Stan to fit the
Thurstonian IRT model in a Bayesian framework.

```{r, results="hide"}
fit <- fit_TIRT_stan(triplets_long, chains = 1)
```

As basic summary and convergence checks can be obtained via

```{r, eval=FALSE}
print(fit)
```

Finally, we obtain predictions of participants' trait scores in a tidy
data format via

```{r}
pr <- predict(fit)
head(pr)
```

The thurstonianIRT package not only comes with model fitting functions but
also with the possibility to simulate data from Thurstonian IRT models.
Below we simulate data with a very similar structure to the `triplets`
data set we have used above.

```{r}
sim_data <- sim_TIRT_data(
  npersons = 200,
  ntraits = 3,
  nblocks_per_trait = 4,
  gamma = 0,
  lambda = runif(12, 0.5, 1),
  Phi = diag(3)
)
head(sim_data)
```

The structure of the data is the same as what we obtain via
the `make_TIRT_data` function and can readily be passed to the model
fitting functions.

## How to install thurstonianIRT

To install the latest release version from CRAN use

```{r, eval = FALSE}
install.packages("thurstonianIRT")
```

The current developmental version can be downloaded from github via

```{r, eval = FALSE}
if (!requireNamespace("remotes")) {
  install.packages("remotes")
}
remotes::install_github("paul-buerkner/thurstonianIRT")
```

## References

Brown, A., & Maydeu-Olivares, A. (2011). Item response modeling of forced-choice questionnaires. *Educational and Psychological Measurement*, 71(3), 460-502.
https://www.doi.org/10.1177/0013164410375112

Brown, A., & Maydeu-Olivares, A. (2012). Fitting a Thurstonian IRT model to forced-choice data using Mplus. *Behavior Research Methods*, 44(4), 1135-1147.
https://www.doi.org/10.3758/s13428-012-0217-x

Bürkner P. C., Schulte N., & Holling H. (2019). On the Statistical and Practical Limitations of Thurstonian IRT Models. *Educational and Psychological Measurement.* https://www.doi.org/10.1177/0013164419832063
